Estimating Access to Rabies Post-Exposure Prophylaxis in Madagascar

### 2019-02-26 accessPEP.Rmd
working version of sensitivity analyses and model validation

```{r directory, echo = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
## Libraries
library(knitr)
library(RImagePalette)
library(jpeg)
library(tidyverse)
library(lubridate)
library(magrittr)
library(rgdal)
library(broom)
library(stringr)
library(ISOweek)
library(raster)
library(sp)
library(MASS)
library(reshape2)
source("R/utils.R")

@@

```{r network output, echo=FALSE, fig.cap="Figure 1A", results = "show", out.width = '100%'}
knitr::include_graphics("figs/network.pdf")
```

@@

``` {r trying to exclude contacts, fig.cap = "Figure 3A"}
## Testing with Moramanga data 
sdev <- seq(2, 10, by = 1)
check <- data.frame(sd = NA, prop_bites = NA, prop_contacts = NA)
for (i in 1:length(sdev)){
  ctar %>%
    group_by(date.reported) %>%
    summarize(n = n()) %>%
    mutate(exclude = ifelse(n > mean(n) + sdev[i]*sd(n), 1, 0)) -> mora_ts
  ctar$exclude <-mora_ts$exclude[match(ctar$date.reported, mora_ts$date.reported)]
  checksd <- table(ctar$consult.type, ctar$exclude)[, 2]/table(ctar$consult.type)
  check <- rbind(check, c(sdev[i], checksd[1:2]))
}

@@

### Models of bites as a function of travel times
```{r likelihood of observing moramanga data @ commune level, fig.cap = "Figure 5A"}

get.likelihood.mora <- function(commune, params) {
  # To test likelihood function
  # commune = comm_data
  # district = exps_dist
  # params = c(1e-6, 0.1)
  B_ttimes <- params[1]
  B_0 <- params[2]
  commune %>%
      mutate(bites_mu = exp(B_ttimes*ttimes_weighted + B_0)*pop) -> sum_bites
  return(-sum(dpois(round(sum_bites$bites), lambda = sum_bites$bites_mu, log = TRUE)))
}

get.likelihood.mada <- function(commune, district, params) {
  # To test likelihood function
  # commune = comm_data
  # district = exps_dist
  # params = c(1e-6, 0.1)
  B_ttimes <- params[1]
  B_0 <- params[2]
  commune %>%
      filter(exclude == 0) %>%
      group_by(mdg_dis_co) %>%
      mutate(bites = exp(B_ttimes*ttimes_weighted + B_0)*pop) %>%
      summarize(sum_bites = sum(bites)) %>%
      left_join(select(district, district, dist_bites = bites), 
                by = c("mdg_dis_co" = "district")) %>%
      filter(!is.na(dist_bites)) -> sum_bites
  return(-sum(dpois(round(sum_bites$dist_bites), lambda = sum_bites$sum_bites, log = TRUE)))
}

get.likelihood.dist <- function(district, params) {
  # To test likelihood function
  # commune = comm_data
  # district = exps_dist
  # params = c(1e-6, 0.1)
  B_ttimes <- params[1]
  B_0 <- params[2]
  district %>%
      filter(exclude == 0) %>%
      mutate(exp_bites = exp(B_ttimes*ttimes_weighted_dist + B_0)*pop) -> district
  return(-sum(dpois(round(district$exp_bites), lambda = round(district$bites), log = TRUE)))
}

@@

## Mada
plot(ttimes_plot/60, exp(B_ttimes_mada*ttimes_plot + B_0_mada)*pop_plot, type = "l", bty = "l",
     ylab = "Mean bites per 100k", xlab = "Travel times (hrs)", col = "darkblue")
points(ttimes_plot/60, exp(B_ttimes_mada_CI[1]*ttimes_plot + B_0_mada_CI[1])*pop_plot, type = "l", 
       lty = 2, col = "lightblue")
points(ttimes_plot/60, exp(B_ttimes_mada_CI[2]*ttimes_plot + B_0_mada_CI[2])*pop_plot, type = "l", 
       lty = 2, col = "lightblue")

points(ttimes_plot/60, exp(B_ttimes_mora*ttimes_plot + B_0_mora)*pop_plot, type = "l", col = "red")
points(ttimes_plot/60, exp(B_ttimes_mora_CI[1]*ttimes_plot + B_0_mora_CI[1])*pop_plot, type = "l", 
       lty = 2, col = "pink")
points(ttimes_plot/60, exp(B_ttimes_mora_CI[2]*ttimes_plot + B_0_mora_CI[2])*pop_plot, type = "l",
       lty = 2, col = "pink")

points(ttimes_plot/60, exp(B_ttimes_dist*ttimes_plot + B_0_dist)*pop_plot, type = "l", col = "purple")
points(ttimes_plot/60, exp(B_ttimes_dist_CI[1]*ttimes_plot + B_0_dist_CI[1])*pop_plot, type = "l", 
       lty = 2, col = "lavender")
points(ttimes_plot/60, exp(B_ttimes_dist_CI[2]*ttimes_plot + B_0_dist_CI[2])*pop_plot, type = "l",
       lty = 2, col = "lavender")

@@

#### Model validation: TO DO

@@

```{r scale example}
## Comparing scales
## Function for reporting
get.reporting_burden <- function(pop, ttimes, B_ttimes, B_0, hdr = 25, incidence = 0.01, 
                          exp_rate = 0.39, p_rabid = 0.6, rho_max = 0.98, p_death = 0.16, 
                          type = "reporting") {
  bites_100k <- exp(B_ttimes*ttimes + B_0)*pop
  R_100k <- pop/hdr*incidence*exp_rate
  if (type == "reporting"){
    rho <- bites_100k*(ifelse(bites_100k*p_rabid > R_100k, rho_max*R_100k/bites_100k, p_rabid))/R_100k
    return(rho)
  } else {
    unreported <- R_100k - bites_100k*(ifelse(bites_100k*p_rabid > R_100k, 
                                              rho_max*R_100k/bites_100k, p_rabid))
    deaths <- 0.16*unreported
    return(deaths)
  }
}

comm_data$reporting_min <- get.reporting_burden(pop = comm_data$pop, ttimes = comm_data$ttimes_weighted, 
                                            B_ttimes = B_ttimes_mada, B_0 = B_0_mada,
                                            hdr = 5, p_rabid = 0.2)

@@

```{r burden results, fig.cap = "Figure 7", cache = TRUE}
## Commune level
source("R/get.burden.R")
results_pr20 <- get.burden(p_rabid = 0.20, rho_max = 0.98)
results_pr60 <- get.burden(p_rabid = 0.60,rho_max = 0.98)
results_pr20$pr <- "0.20"
results_pr60$pr <- "0.60"
results <- rbind(results_pr20, results_pr60)

results %>% 
  group_by(mdg_dis_co, pr) %>% 
  summarize(tot = sum(deaths_mean), upperCI = sum(deaths_upperCI),
            lowerCI = sum(deaths_lowerCI)) %>%
  left_join(dist_catchments) %>% 
  left_join(select(mada_district@data, mdg_dis_co, district)) -> results_grouped

## District level
results_pr20_dist <- get.burden.dist(p_rabid = 0.20, rho_max = 0.98)
results_pr60_dist <- get.burden.dist(p_rabid = 0.60,rho_max = 0.98)
results_pr20_dist$pr <- "0.20"
results_pr60_dist$pr <- "0.60"
results_dist <- rbind(results_pr20_dist, results_pr60_dist)

results_dist %>% 
  left_join(results_grouped, by = c("mdg_dis_co" = "mdg_dis_co", "pr" = "pr")) -> results_comp

ggplot(results_comp, aes (x = deaths_mean, y = tot, color = ttimes_weighted.x)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "grey", linetype = "dashed") +
  facet_wrap (~pr)

results_dist %>% 
  group_by(mdg_dis_co, pr) %>% 
  summarize(tot = sum(deaths_mean), upperCI = sum(deaths_upperCI),
            lowerCI = sum(deaths_lowerCI)) %>%
  left_join(dist_catchments) %>% 
  left_join(select(mada_district@data, mdg_dis_co, district)) -> results_grouped_dist
results_grouped_dist$scale <- "district"
results_grouped$scale <- "commune"

results_all <- rbind(results_grouped, results_grouped_dist)

ggplot(results_all, aes(y = reorder(district, ttimes_weighted), x = tot, 
                              color = pr)) +
  ggstance::geom_pointrangeh(aes(xmin = lowerCI, xmax = upperCI), fatten = 2) +
  ylab("") +
  xlab("Average annual deaths") +
  scale_color_manual(values = c("0.20" = alpha("red", 0.2), "0.60" = alpha("red", 0.6)), 
                     name = "Proportion rabid") +
  facet_wrap(~scale) -> p
p + theme(axis.text.y = element_text(margin = margin(t = 10, r = 0, b = 10, l = 0), size = 7))

## Maps of burden
## district level
mada_district %>%
  tidy(., region = "mdg_dis_co") %>%
  left_join(results_pr60_dist, by = c("id" = "mdg_dis_co")) -> ggmada_dist

ggplot(ggmada_dist, aes(long, lat, group = group, fill = deaths_mean)) +
  geom_polygon() +
  scale_fill_continuous(type = "viridis") -> burden_map
burden_map
ggsave("burden_map_district.pdf")

## commune level
mada_communes %>%
  tidy(., region = "mdg_cm_") %>%
  left_join(results_pr60, by = c("id" = "mdg_cm_")) -> ggmada_commune

ggplot(ggmada_commune, aes(long, lat, group = group, fill = deaths_mean)) +
  geom_polygon() +
  scale_fill_continuous(type = "viridis") -> burden_map
burden_map
ggsave("burden_map_commune.pdf")

@@

When we estimate burden of deaths stochastically within this range of incidence and given our high and low estimates of proportion of reported bites that are rabid, we see that burden of deaths also decreases with travel times (Figure 7, results presented aggregated at the district level). Overall, we estimate average annual deaths between `r round(sum(results_pr60$deaths_mean))` - `r round(sum(results_pr20$deaths_mean))`. **Also estimate deaths averted here!**.

>>>>

```{r directory, echo = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
## Libraries
library(data.table)
library(knitr)
library(RImagePalette)
library(jpeg)
library(tidyverse)
library(lubridate)
library(magrittr)
library(rgdal)
library(broom)
library(stringr)
library(ISOweek)
library(raster)
library(sp)
library(MASS)
library(reshape2)
source("R/utils.R")

@@

```{r network output, echo=FALSE, fig.cap="Figure 1A", results = "show", out.width = '100%'}
knitr::include_graphics("figs/network.jpeg")
```

@@

```{r trying to exclude contacts, fig.cap="Figure 3A"}
## Testing with Moramanga data 
sdev <- seq(2, 10, by = 1)
check <- data.frame(sd = NA, prop_bites = NA, prop_contacts = NA)
for (i in 1:length(sdev)){
  ctar %>%
    group_by(date.reported) %>%
    summarize(n = n()) %>%
    mutate(exclude = ifelse(n > mean(n) + sdev[i]*sd(n), 1, 0)) -> mora_ts
  ctar$exclude <-mora_ts$exclude[match(ctar$date.reported, mora_ts$date.reported)]
  checksd <- table(ctar$consult.type, ctar$exclude)[, 2]/table(ctar$consult.type)
  check <- rbind(check, c(sdev[i], checksd[1:2]))
}

@@

### Models of bites as a function of travel times
```{r likelihood of observing moramanga data @ commune level, fig.cap = "Figure 5A"}
library(boot)

get.likelihood.bites <- function(bites, covar, pop, params) {
  beta <- params[1]
  intercept <- params[2]
  exp_bites <- inv.logit(beta*covar + intercept)*pop
  return(-sum(dpois(round(bites), lambda = exp_bites, log = TRUE)))
}

get.likelihood.mora <- function(commune, params) {
  # To test likelihood function
  # commune = comm_data
  # district = exps_dist
  # params = c(1e-6, 0.1)
  B_ttimes <- params[1]
  B_0 <- params[2]
  commune %>%
      mutate(bites_mu = inv.logit(B_ttimes*ttimes_weighted + B_0)*pop) -> sum_bites
  return(-sum(dpois(round(sum_bites$), lambda = sum_bites$bites_mu, log = TRUE)))
}

get.likelihood.mada <- function(commune, district, params) {
  # To test likelihood function
  # commune = comm_data
  # district = exps_dist
  # params = c(1e-6, 0.1)
  B_ttimes <- params[1]
  B_0 <- params[2]
  commune %>%
      filter(exclude == 0) %>%
      group_by(mdg_dis_co) %>%
      mutate(bites = inv.logit(B_ttimes*ttimes_weighted + B_0)*pop) %>%
      summarize(sum_bites = sum(bites)) %>%
      left_join(select(district, district, dist_bites = bites), 
                by = c("mdg_dis_co" = "district")) %>%
      filter(!is.na(dist_bites)) -> sum_bites
  return(-sum(dpois(round(sum_bites$dist_bites), lambda = sum_bites$sum_bites, log = TRUE)))
}
get.likelihood.dist <- function(district, params) {
  # To test likelihood function
  # commune = comm_data
  # district = exps_dist
  # params = c(1e-6, 0.1)
  B_ttimes <- params[1]
  B_0 <- params[2]
  district %>%
      filter(exclude == 0) %>%
      mutate(exp_bites = inv.logit(B_ttimes*ttimes_weighted_dist + B_0)*pop) -> district
  return(-sum(dpois(round(district$exp_bites), lambda = round(district$bites), log = TRUE)))
}

@@

## Mada
par_df <- as.data.frame(list(model = c("Mada communes", "Moramanga communes", "Mada districts"),
                     B_ttimes = c(B_ttimes_mada, B_ttimes_mora, B_ttimes_dist),
                     B_0 = c(B_0_mada, B_0_mora, B_0_dist),
                     B_0_upper = c(B_0_mada_CI[2], B_0_mora_CI[2], B_0_dist_CI[2]),
                     B_0_lower = c(B_0_mada_CI[1], B_0_mora_CI[1], B_0_dist_CI[1]),
                     B_ttimes_upper = c(CI = B_ttimes_mada_CI[2], B_ttimes_mora_CI[2],
                                  B_ttimes_dist_CI[2]),
                     B_ttimes_lower = c(CI = B_ttimes_mada_CI[1], B_ttimes_mora_CI[1],
                                  B_ttimes_dist_CI[1])))

preds <- foreach(i = 1:nrow(par_df), .combine = rbind) %do% {
  bites <- inv.logit(par_df$B_ttimes[i]*ttimes_plot + par_df$B_0[i])*pop_plot
  upper <- inv.logit(par_df$B_ttimes_upper[i]*ttimes_plot + par_df$B_0_upper[i])*pop_plot
  lower <- inv.logit(par_df$B_ttimes_lower[i]*ttimes_plot + par_df$B_0_lower[i])*pop_plot
  return(as.data.frame(list(bites = bites, upper = upper, lower = lower, ttimes = ttimes_plot/60, 
                            model = par_df$model[i])))
}
ggplot(data = preds, aes(x = ttimes, y = bites, color = model)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = model), alpha = 0.5) +

plot(ttimes_plot/60, inv.logit(B_ttimes_mada*ttimes_plot + B_0_mada)*pop_plot, type = "l", bty = "l",
     ylab = "Mean bites per 100k", xlab = "Travel times (hrs)", col = "darkblue")
points(ttimes_plot/60, inv.logit(B_ttimes_mada_CI[1]*ttimes_plot + B_0_mada_CI[1])*pop_plot, type = "l", 
       lty = 2, col = "lightblue")
points(ttimes_plot/60, inv.logit(B_ttimes_mada_CI[2]*ttimes_plot + B_0_mada_CI[2])*pop_plot, type = "l", 
       lty = 2, col = "lightblue")

points(ttimes_plot/60, inv.logit(B_ttimes_mora*ttimes_plot + B_0_mora)*pop_plot, type = "l", col = "red")
points(ttimes_plot/60, inv.logit(B_ttimes_mora_CI[1]*ttimes_plot + B_0_mora_CI[1])*pop_plot, type = "l", 
       lty = 2, col = "pink")
points(ttimes_plot/60, inv.logit(B_ttimes_mora_CI[2]*ttimes_plot + B_0_mora_CI[2])*pop_plot, type = "l",
       lty = 2, col = "pink")

points(ttimes_plot/60, inv.logit(B_ttimes_dist*ttimes_plot + B_0_dist)*pop_plot, type = "l", col = "purple")
points(ttimes_plot/60, inv(B_ttimes_dist_CI[1]*ttimes_plot + B_0_dist_CI[1])*pop_plot, type = "l", 
       lty = 2, col = "lavender")
points(ttimes_plot/60, exp(B_ttimes_dist_CI[2]*ttimes_plot + B_0_dist_CI[2])*pop_plot, type = "l",
       lty = 2, col = "lavender")

@@

#### Model validation
```{r model validation}
## Models
get.likelihood.other <- function(weight, bites, pop, params) {
  B_ttimes <- params[1]
  B_0 <- params[2]
  exp_bites <- inv.logit(B_ttimes*weight + B_0)*pop
  return(-sum(dpois(round(exp_bites), lambda = round(bites), log = TRUE)))
}

## Distance to CTAR
ctar_coords <- cbind(ctar_metadata$LONGITUDE, ctar_metadata$LATITUDE)
district_coords <- coordinates(mada_district)
library(geosphere)
distance_mat <- distm(district_coords, ctar_coords)
dist_min <- apply(distance_mat, 1, min)/1000
exps_dist$dist_min <- dist_min[match(exps_dist$district, mada_district$mdg_dis_co)]

## Unweighted travel times
exps_dist$ttimes_unweighted <- mada_district$study_area[match(exps_dist$district,
                                                              mada_district$mdg_dis_co)]

### District parameters
dist_pars <- optim(par = c(1e-5, -0.1), get.likelihood.other, 
                   weight = exps_dist$dist_min, bites = exps_dist$bites, pop = exps_dist$pop,
                   hessian = TRUE)
B_ttimes_dist <- dist_pars$par[1]
B_0_dist <- dist_pars$par[2]

plot(exps_dist$bites, inv.logit(B_ttimes_dist*exps_dist$dist_min+ B_0_dist)*exps_dist$pop)
plot(exps_dist$ttimes_weighted_dist, exps_dist$bites)

### Unweighted
dist_pars <- optim(par = c(1e-3, 0.1), get.likelihood.other, 
                   weight = exps_dist$ttimes_unweighted, bites = exps_dist$bites, 
                   pop = exps_dist$pop,
                   hessian = TRUE)
B_ttimes_dist <- dist_pars$par[1]
B_0_dist <- dist_pars$par[2]

plot(exps_dist$bites, exp(B_ttimes_dist*exps_dist$ttimes_unweighted+ B_0_dist))
plot(exps_dist$ttimes_weighted_dist, exps_dist$dist_min)

### Pop
get.likelihood.pop <- function(bites, pop, params) {
  B_ttimes <- params[1]
  B_0 <- params[2]
  exp_bites <- exp(B_ttimes*pop/1e5 + B_0)
  return(-sum(dpois(round(exp_bites), lambda = round(bites), log = TRUE)))
}
pop_pars <- optim(par = c(1e-5, 0.1), get.likelihood.pop, 
                   pop = exps_dist$pop, bites = exps_dist$bites, 
                   hessian = TRUE)
B_pop <- pop_pars$par[1]
B_0_pop <- pop_pars$par[2]
check <- glm(round(bites) ~ pop, data = exps_dist, family = "poisson")
plot(exps_dist$bites, exp(B_pop*exps_dist$pop/1e5 + B_0_pop), pch = 20, col = alpha("red", 0.5))
plot(exps_dist$bites, predict(check, type = "response"), pch = 20, col = alpha("orange", 0.5))
abline(a = 0, b = 1, col = "grey", lty = 2)

## Getting standard errors
# this is how you can get standard errors from optim, but inverting the hessian
se <- sqrt(abs(diag(solve(pop_pars$hessian)))); se

#simulate "standard errors"
varcov <- solve(pop_pars$hessian)

ntest = 1000; testpar = pop_pars$par; sim.par = matrix(NA, ntest, 2)
for (k in 1:ntest) {
  sim.par[k, ] = mvrnorm(1, pop_pars$par, varcov)
}

B_pop_CI <- quantile(sim.par[,1],c(0.025,0.975))
B_0_pop_CI <- quantile(sim.par[,2],c(0.025,0.975))

## @ commune level
get.likelihood.mada <- function(commune, district, params) {
  # To test likelihood function
  # commune = comm_data
  # district = exps_dist
  # params = c(1e-6, 0.1)
  B_ttimes <- params[1]
  B_0 <- params[2]
  commune %>%
      filter(exclude == 0) %>%
      group_by(mdg_dis_co) %>%
      mutate(bites = exp(B_ttimes*pop/1e5 + B_0)) %>%
      summarize(sum_bites = sum(bites)) %>%
      left_join(select(district, district, dist_bites = bites), 
                by = c("mdg_dis_co" = "district")) %>%
      filter(!is.na(dist_bites)) -> sum_bites
  return(-sum(dpois(round(sum_bites$dist_bites), lambda = sum_bites$sum_bites, log = TRUE)))
}
comm_pars <- optim(par = c(1e-3, 0.1), get.likelihood.mada, 
                   commune = comm_data, district = exps_dist, 
                   hessian = TRUE)
B_ttimes_comm <- comm_pars$par[1]
B_0_comm <- comm_pars$par[2]

## Getting standard errors
# this is how you can get standard errors from optim, but inverting the hessian
se <- sqrt(abs(diag(solve(comm_pars$hessian)))); se

#simulate "standard errors"
varcov <- solve(comm_pars$hessian)

ntest = 1000; testpar = comm_pars$par; sim.par = matrix(NA, ntest, 2)
for (k in 1:ntest) {
  sim.par[k, ] = mvrnorm(1, comm_pars$par, varcov)
}

B_ttimes_comm_CI <- quantile(sim.par[,1],c(0.025,0.975))
B_0_comm_CI <- quantile(sim.par[,2],c(0.025,0.975))

comm_data %>%
      filter(exclude == 0) %>%
      mutate(bites = exp(B_ttimes_comm*pop/1e5 + B_0_comm)) %>%
      group_by(mdg_dis_co) %>%
      summarize(sum_bites = sum(bites)) %>%
      left_join(select(exps_dist, district, dist_bites = bites), 
                by = c("mdg_dis_co" = "district")) %>%
      filter(!is.na(dist_bites)) -> sum_bites
sum_bites %$% plot(dist_bites, sum_bites, bty = "n", pch = 20, col = alpha("darkblue", 0.5),
                   xlab = "Observed mean bites", ylab = "Model predicted mean bites", 
                   xlim = c(0, 5000))
comm_data %>%
      filter(exclude == 0) %>%
      mutate(bites = exp(B_ttimes_mada*ttimes_weighted + B_0_mada)*pop) %>%
      group_by(mdg_dis_co) %>%
      summarize(sum_bites = sum(bites)) %>%
      left_join(select(exps_dist, district, dist_bites = bites), 
                by = c("mdg_dis_co" = "district")) %>%
      filter(!is.na(dist_bites)) -> sum_bites
sum_bites %$% plot(dist_bites, sum_bites, pch = 20, col = alpha("red", 0.5))
abline(a = 0, b = 1, col = "grey", lty = 2)

exps_dist %>%
      mutate(bites_mu = exp(B_ttimes_dist*ttimes_weighted_dist + B_0_dist)*pop) -> check1
points(check1$bites, check1$bites_mu, col = alpha("purple", 0.5), pch = 20)
points(exps_dist$bites, predict(check, type = "response"), pch = 20, col = alpha("orange", 0.5))
abline(a = 0, b = 1, col = "grey", lty = 2)

mora_bites %>%
      mutate(bites_mu = exp(B_ttimes_mora*ttimes_weighted + B_0_mora)*pop) -> check
points(check$bites, check$bites_mu, bty = "n", pch = 20, col = alpha("blue", 0.5))
check <- glm(round(bites) ~ pop, data = mora_bites, family = "poisson")
plot(mora_bites$bites, predict(check, type = "response"), pch = 20, col = alpha("orange", 0.5))
abline(a = 0, b = 1, col = "grey", lty = 2)

## Correlation plot

## Table of fits/parameter results

## Impact of contacts

## Impact of RLE threshold

@@

```{r scale example}
## Comparing scales
comm_data$reporting_min <- get.reporting_burden(pop = comm_data$pop, ttimes = comm_data$ttimes_weighted, 
                                            B_ttimes = B_ttimes_mada, B_0 = B_0_mada,
                                            hdr = 5, p_rabid = 0.2)

@@

```{r burden results, fig.cap = "Figure 7", cache = TRUE}
# Incremental analysis ------------------------------------------------------------------------
## District travel times
dist_ttimes <- read.csv("output/district_temp_scenario_20190212_181056.csv", row.names = 1)
comm_ttimes <- read.csv("output/commune_temp_scenario_20190212_182810.csv", row.names = 1)

dist_ttimes <- cbind(dist_catchments$ttimes_weighted, dist_ttimes)
dist_mat <- matrix(NA, nrow(dist_ttimes), ncol(dist_ttimes))
dist_mat[, 1] <- dist_ttimes[, 1]
for (i in 2:ncol(dist_ttimes)){
  dist_mat[, i] <- ifelse(dist_ttimes[, i] < dist_ttimes[, i - 1], dist_ttimes[, i], NA)
}
dist_mat <- as.data.table(dist_mat)
dist_mat$mdg_dis_co <- as.character(mada_district$mdg_dis_co)
dist_mat <- melt(dist_mat, id.vars = "mdg_dis_co", value.name = "ttimes", variable.name = "scenario")
dist_mat$scenario <- as.numeric(gsub("V", "", dist_mat$scenario)) - 1

dist_mat %>%
  drop_na(ttimes) %>%
  left_join(select(mada_district@data, mdg_dis_co, district, pop2015adj)) -> dist_mat

## Commune travel times
comm_ttimes <- cbind(comm_catchments$ttimes_weighted, comm_ttimes)
comm_mat <- matrix(NA, nrow(comm_ttimes), ncol(comm_ttimes))
comm_mat[, 1] <- comm_ttimes[, 1]
for (i in 2:ncol(comm_ttimes)){
  comm_mat[, i] <- ifelse(comm_ttimes[, i] < comm_ttimes[, i - 1], comm_ttimes[, i], NA)
}
comm_mat <- as.data.table(comm_mat)
comm_mat$mdg_cm_ <- mada_communes$mdg_cm_
comm_mat <- melt(comm_mat, id.vars = "mdg_cm_", value.name = "ttimes", variable.name = "scenario")
comm_mat$scenario <- as.numeric(gsub("V", "", comm_mat$scenario)) - 1

comm_mat %>%
  drop_na(ttimes) %>%
  left_join(select(comm_data, mdg_cm_, pop)) -> comm_mat

## Estimate burden
source("R/get.burden.R")
system.time({
  results_pr20_comm <- get.burden(names = comm_mat$mdg_cm_, ttimes = comm_mat$ttimes,
                           pop = comm_mat$pop, scenario = comm_mat$scenario,
                           param_ttimes = B_ttimes_mada,
                           param_intercept = B_0_mada,
                           p_rabid = 0.2, rho_max = 0.98,
                           max_HDR = 25, min_HDR = 5, 
                           dog_rabies_inc = 0.01, human_exp_rate = 0.39, 
                           prob_death = 0.16, nsims = 1000)

results_pr60_comm <- get.burden(names = comm_mat$mdg_cm_, ttimes = comm_mat$ttimes,
                           pop = comm_mat$pop, scenario = comm_mat$scenario,
                           param_ttimes = B_ttimes_mada,
                           param_intercept = B_0_mada,
                           p_rabid = 0.6, rho_max = 0.98,
                           max_HDR = 25, min_HDR = 5, 
                           dog_rabies_inc = 0.01, human_exp_rate = 0.39, 
                           prob_death = 0.16, nsims = 1000)
})

## Filling it in
results_pr20_comm %>%
  complete(scenario, names) %>%
  group_by(names) %>%
  arrange(scenario) %>%
  fill(3:ncol(.), .direction = "down") -> results_pr20_comm

results_pr60_comm %>%
  complete(scenario, names) %>%
  group_by(names) %>%
  arrange(scenario) %>%
  fill(3:ncol(.), .direction = "down") -> results_pr60_comm

system.time({
results_pr20_dist <- get.burden(names = dist_mat$mdg_dis_co, ttimes = dist_mat$ttimes,
                           pop = dist_mat$pop, scenario = dist_mat$scenario,
                           param_ttimes = B_ttimes_dist,
                           param_intercept = B_0_dist,
                           p_rabid = 0.2, rho_max = 0.98,
                           max_HDR = 25, min_HDR = 5, 
                           dog_rabies_inc = 0.01, human_exp_rate = 0.39, 
                           prob_death = 0.16, nsims = 1000)

results_pr60_dist <- get.burden(names = dist_mat$mdg_dis_co, ttimes = dist_mat$ttimes,
                           pop = dist_mat$pop, scenario = dist_mat$scenario,
                           param_ttimes = B_ttimes_dist,
                           param_intercept = B_0_dist,
                           p_rabid = 0.6, rho_max = 0.98,
                           max_HDR = 25, min_HDR = 5, 
                           dog_rabies_inc = 0.01, human_exp_rate = 0.39, 
                           prob_death = 0.16, nsims = 1000)
})

results_pr20_dist %>%
  complete(scenario, names) %>%
  group_by(names) %>%
  arrange(scenario) %>%
  fill(3:ncol(.), .direction = "down") -> results_pr20_dist

results_pr60_dist %>%
  complete(scenario, names) %>%
  group_by(names) %>%
  arrange(scenario) %>%
  fill(3:ncol(.), .direction = "down") -> results_pr60_dist
```

``` {r plotting burden} 
## Group commune results to district level
results_comm <- rbind(results_pr20_comm, results_pr60_comm)
colnames(results_comm) <- gsub("results.", "", colnames(results_comm))
results_comm %>%
  left_join(select(comm_data, mdg_dis_co, mdg_cm_), 
            by = c("names" = "mdg_cm_")) %>%
  group_by(scenario, mdg_dis_co, pr) %>%
  summarize(names = first(mdg_dis_co), 
            deaths_mean = sum(deaths_mean, na.rm = TRUE), 
            deaths_upperCI = sum(deaths_upperCI, na.rm = TRUE),
            deaths_lowerCI = sum(deaths_lowerCI, na.rm = TRUE)) -> results_comm_summarized
results_comm_summarized$model <- "commune"
results_dist <- rbind(results_pr20_dist, results_pr60_dist)
colnames(results_dist) <- gsub("results.", "", colnames(results_dist))
results_dist %>%
  mutate(model = "district") %>%
  select(scenario, names, pr, deaths_mean, deaths_lowerCI, deaths_upperCI, 
         model) %>%
  bind_rows(., results_comm_summarized) -> results_all
results_all %>%
  left_join(select(mada_district@data, district, mdg_dis_co, ttimes_weighted),
            by = c("names" = "mdg_dis_co")) -> results_all

## Plot national level burden
results_all %>% 
  group_by(scenario, pr, model) %>% 
  summarize(deaths_mean = sum(deaths_mean, na.rm = TRUE), 
            deaths_upperCI = sum(deaths_upperCI, na.rm = TRUE),
            deaths_lowerCI = sum(deaths_lowerCI, na.rm = TRUE)) -> results_natl

## Plotting national level
ggplot(data = results_natl, aes (x = scenario, y = deaths_mean,
                                 color = interaction(model, pr))) +
  geom_line() +
  geom_ribbon(aes(ymin = deaths_upperCI, ymax = deaths_lowerCI, color = NULL, 
                  fill = interaction(model, pr)), alpha = 0.5) +
  scale_color_manual(values = c("district.0.2" = "blue", 
                                "district.0.6" = "darkblue", 
                                "commune.0.2" = "magenta", 
                                "commune.0.6" = "purple"), 
                     name = "model pars") +
  scale_fill_manual(values = c("district.0.2" = "blue", 
                                "district.0.6" = "darkblue", 
                                "commune.0.2" = "magenta", 
                                "commune.0.6" = "purple"), 
                    guide = "none")

## Plotting national level
ggplot(data = results_all, aes(x = deaths_mean, y = reorder(district, ttimes_weighted), 
                               color = scenario, shape = as.factor(pr))) +
  ggstance::geom_pointrangeh(aes(xmin = deaths_lowerCI, xmax = deaths_upperCI), alpha = 0.25,
                             fatten = 2) +
  facet_wrap(~model + as.factor(pr))

## Baseline results
results_all %>% filter(scenario == 0) -> results_all_base
ggplot(results_all_base, aes(y = reorder(district, ttimes_weighted), x = deaths_mean, 
                              color = model, shape = as.factor(pr))) +
  ggstance::geom_pointrangeh(aes(xmin = deaths_lowerCI, xmax = deaths_upperCI), fatten = 2) +
  ylab("") +
  xlab("Average annual deaths") +
  scale_color_manual(values = c("district" = alpha("blue", 0.5), 
                                "commune" = alpha("purple", 0.5)), 
                     name = "Scale") +
  facet_wrap(~ as.factor(pr)) -> p
p + theme(axis.text.y = element_text(margin = margin(t = 10, r = 0, b = 10, l = 0), size = 7))

@@

```{r doing it from the median, cache=TRUE}
system.time({
  results_comm_med <- get.burden_med(names = comm_mat$mdg_cm_, ttimes = comm_mat$ttimes,
                           pop = comm_mat$pop, scenario = comm_mat$scenario,
                           param_ttimes = B_ttimes_mada,
                           param_intercept = B_0_mada,
                           p_rab_min = 0.2, p_rab_max = 0.6, rho_max = 0.98,
                           max_HDR = 25, min_HDR = 5, 
                           dog_rabies_inc = 0.01, human_exp_rate = 0.39, 
                           prob_death = 0.16, nsims = 1000)
})

## Filling it in
results_comm_med %>%
  complete(scenario, names) %>%
  group_by(names) %>%
  arrange(scenario) %>%
  fill(3:ncol(results_comm_med), .direction = "down") -> results_comm_med

system.time({
results_dist_med <- get.burden_med(names = dist_mat$mdg_dis_co, ttimes = dist_mat$ttimes,
                           pop = dist_mat$pop, scenario = dist_mat$scenario,
                           param_ttimes = B_ttimes_dist,
                           param_intercept = B_0_dist,
                           p_rab_min = 0.2, p_rab_max = 0.6, rho_max = 0.98,
                           max_HDR = 25, min_HDR = 5, 
                           dog_rabies_inc = 0.01, human_exp_rate = 0.39, 
                           prob_death = 0.16, nsims = 1000)
})

results_dist_med %>%
  complete(scenario, names) %>%
  group_by(names) %>%
  arrange(scenario) %>%
  fill(3:ncol(.), .direction = "down") -> results_dist_med

## Group commune results to district level
colnames(results_comm_med) <- gsub("results.", "", colnames(results_comm_med))
results_comm_med %>%
  left_join(select(comm_data, mdg_dis_co, mdg_cm_), 
            by = c("names" = "mdg_cm_")) %>%
  group_by(scenario, mdg_dis_co) %>%
  summarize(deaths_mean = sum(deaths_mean, na.rm = TRUE), 
            deaths_upperCI = sum(deaths_upperCI, na.rm = TRUE),
            deaths_lowerCI = sum(deaths_lowerCI, na.rm = TRUE)) %>%
  rename(names = mdg_dis_co) -> results_comm_summarized
results_comm_summarized$model <- "commune"
colnames(results_dist_med) <- gsub("results.", "", colnames(results_dist_med))
results_dist_med %>%
  mutate(model = "district") %>%
  select(scenario, names, deaths_mean, deaths_lowerCI, deaths_upperCI, 
         model) -> results_dist_med
results_all <- bind_rows(results_dist_med, results_comm_summarized)
results_all %>%
  left_join(select(mada_district@data, district, mdg_dis_co, ttimes_weighted),
            by = c("names" = "mdg_dis_co")) -> results_all

## Plot national level burden
results_all %>% 
  group_by(scenario, model) %>% 
  summarize(deaths_mean = sum(deaths_mean, na.rm = TRUE), 
            deaths_upperCI = sum(deaths_upperCI, na.rm = TRUE),
            deaths_lowerCI = sum(deaths_lowerCI, na.rm = TRUE)) -> results_natl
